Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-09-13 snapshot_system/fast14_background.tex

\section{Background and Design Considerations}
\label{sect:option}
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.  
As discussed earlier, collocating a backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architecture where
each physical machine runs a backup service and a distributed file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshots.

@@

Our key design consideration is VM dependence minimization during deduplication and file block management.
\begin{itemize}
\item {\em Deduplication localization.}
Because a data chunk is compared with signatures collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependencies among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly. 
The mark-sweep approach~\cite{mark-sweep} is effective for deletion, and its main cost
is to count if a data chunk is still shared by other snapshots.
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can faciltate parallel execition of snaptshot operations.
\item{\em  Management of file system blocks.}
The file block size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
 a data chunk in a typical deduplication system is of a nonuniform size with 4KB or 8KB on average.
Packaging data chunks to a file system block can create more data dependence among VMs
since a file block can be shared with even more VMs.
Thus we need to consider a minimum association of a fie system block to VMs in the packaging process.
\end{itemize}

@@

We call the traditional deduplication approach as   VM-oblivious (VO)
because they compare fingerprints of snapshots without consideration of VM.
With the above  considerations in mind, we study a 
VM-centric approach (called VC)
for a collocated backup service with resource usage friendly
to the existing applications.  
In designing a VC duplication algorithm, we have considered and adopted some of
the following previously-developed techniques.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized. 
We are seeking for a tradeoff since 
cross-VM  signature comparison can deliver additional compression~\cite{Guo2011,Dong,ExtremeBining}.
\item {\bf Stateless  Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing~\cite{Dong20??}
that divides the deduplication work with an approximation. This work 
is similar to Exreme Binning\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
\item {\bf Sampled Index}.
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplicate request
may access a remote machine for index lookup and the overall overhead of latency latency for all requests
can be significant.  
\end{itemize}

>>>>

\section{Background and Design Considerations}
\label{sect:background}
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.  
As discussed earlier, collocating a backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architecture where
each physical machine runs a backup service and a distributed file system (DFS)~\cite{googlefs03,hdfs10} 
serves a backup store  for the snapshots.

@@

Our key design consideration is VM dependence minimization during deduplication and file block management.
\begin{itemize}
\item {\em Deduplication localization.}
Because a data chunk is compared with signatures collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependencies among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly. 
The mark-sweep approach~\cite{Guo2011} is effective for deletion, and its main cost
is to count if a data chunk is still shared by other snapshots.
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can faciltate parallel execition of snaptshot operations.
\item{\em  Management of file system blocks.}
The file block size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
 a data chunk in a typical deduplication system is of a nonuniform size with 4KB or 8KB on average.
Packaging data chunks to a file system block can create more data dependence among VMs
since a file block can be shared with even more VMs.
Thus we need to consider a minimum association of a fie system block to VMs in the packaging process.
\end{itemize}

@@

We call the traditional deduplication approach as   VM-oblivious (VO)
because they compare fingerprints of snapshots without consideration of VM.
With the above  considerations in mind, we study a 
VM-centric approach (called VC)
for a collocated backup service with resource usage friendly
to the existing applications.  
In designing a VC duplication algorithm, we have considered and adopted some of
the following previously-developed techniques.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized. 
We are seeking for a tradeoff since 
cross-VM  signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
\item {\bf Stateless  Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing~\cite{Dong2011}
that divides the deduplication work with an approximation. This work 
is similar to Exreme Binning\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
\item {\bf Sampled Index}.
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplicate request
may access a remote machine for index lookup and the overall overhead of latency latency for all requests
can be significant.  
\end{itemize}

### 2013-09-13 snapshot_system/fast14_introduction.tex

%************** Paper sections summary
%THIS NEEDS MODIFICATION
The rest of this paper is organized as follows.
Section~\ref{option} reviews the background and discusses the  design options for snapshot backup 
with a VM-centric approach. 
Section~\ref{sect:model}  analyzes the tradeoffs and benefits of our approach. 
Section~\ref{sect:arch}  describes our system architecture and implementation details.
%   the benefit of our approach for fault isolation. 
Section~\ref{exper} is our experimental evaluation that compares with other approaches.
Section~\ref{conc}  concludes this paper.

>>>>

%************** Paper sections summary
%THIS NEEDS MODIFICATION
The rest of this paper is organized as follows.
Section~\ref{sect:background} reviews the background and discusses the  design options for snapshot backup 
with a VM-centric approach. 
Section~\ref{sect:deduplication}  analyzes the tradeoffs and benefits of our approach. 
Section~\ref{sect:architecture}  describes our system architecture and implementation details.
%   the benefit of our approach for fault isolation. 
Section~\ref{sect:evaluation} is our experimental evaluation that compares with other approaches.
Section~\ref{sect:conclusion}  concludes this paper.

### 2013-09-13 snapshot_system/impl.tex

\begin{enumerate}
\item {\bf Creating bloom filter} Scan all the living snapshot recipess and their segment recipes,
for every reference pointing to append store, add it to the bloom filter.
\item {\bf Check existance} For every data reference in the deleted snapshot recipe and its segment recipes,
check the existance of that data reference in bloom filter. If not found, it is safe to delete that piece of data from append store
because no living snapshots has referenced it.
\end{enumerate}

The overall time of running a approximate deletion for one snapshot deletion would be scanning
all the living snapshots and deleted snapshots, since operations on the in-memory bloom filter can be done in
parallel and is much faster than loading recipes from DFS:
\begin{equation}
T = (N_{SS} + 1) * T_{scan\_recipes}
\end{equation}

Using the example and analysis in previous section, this approximate deletion can be done in 5 minutes. 
Memory usage of the bloom filter depends on its false-positive probility $P_{bl}$,
when set $P_{bl}$ to 0.01, the memory footprint of approximate deletion is about 15 MB.

>>>>

\begin{enumerate}
\item {\bf Creating Bloom filter} Scan all the living snapshot recipess and their segment recipes,
for every reference pointing to append store, add it to the Bloom filter.
\item {\bf Check existance} For every data reference in the deleted snapshot recipe and its segment recipes,
check the existance of that data reference in Bloom filter. If not found, it is safe to delete that piece of data from append store
because no living snapshots has referenced it.
\end{enumerate}

The overall time of running a approximate deletion for one snapshot deletion would be scanning
all the living snapshots and deleted snapshots, since operations on the in-memory Bloom filter can be done in
parallel and is much faster than loading recipes from DFS:
\begin{equation}
T = (N_{SS} + 1) * T_{scan\_recipes}
\end{equation}

Using the example and analysis in previous section, this approximate deletion can be done in 5 minutes. 
Memory usage of the Bloom filter depends on its false-positive probility $P_{bl}$,
when set $P_{bl}$ to 0.01, the memory footprint of approximate deletion is about 15 MB.
